---
title: "Homework 1"
author: Zaynub Ibrahim 
output: github_document
---

Solution to HW1. 

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Problem 1

Create a data frame with the specified elements 

```{r create_df}
prob1_df =
  tibble(
   samp = rnorm(10),
   samp_gt_0 = samp > 0, 
   char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
   factor_vec =  factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
 )
```

Take the mean of each variable in my data frame 

```{r mean_pull_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

I can take the mean of numbers and logical but not character or factor because they are not numerical values. 

```{r eval = FALSE}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

The as.numeric function can be applied to numbers, logical, and factors but not character. The factor variable has levels which allows it to work with as.numeric. 

```{r}
as.numeric(pull(prob1_df, samp_gt_0))  * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0))  * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```

The result from trying to convert logical to factor and then multiplying the random sample by the result is not meaningful because is a factor does not have numerical value and cannot be multiplied. 

## Problem 2

```{r}
data("penguins", package = "palmerpenguins")
summary(penguins)
view(penguins)
nrow(penguins)
ncol(penguins)
```

The data in the penguins data set includes the following variables: species, island, bill length, bill depth, flipper length, body mass, sex, and year. 

The species of interest are Adelie, Chinstrap, and Gentoo. 
The islands of interest are Biscoe, Dream, and Torgersen.

This data set contains 344 observations, with `r nrow(penguins)` rows and `r ncol(penguins)` columns. 

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`. 

```{r  scatter_plot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_scatter_plot.pdf")
```



